Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm

### 2011-07-21 ltx/data_paper/DataPaper.tex
minor edits

Here we describe the problem of estimating non-linear state space neural field model using expectation-maximisation (EM) based algorithm \cite{Dempster1977,Shumway2000}. EM is an iterative algorithm that converges on the maximum-likelihood estimates of the parameters. Both the E and the M steps of the EM algorithm increase a lower bound on the log-likelihood function of the parameters given the observed field.  In the E-step we estimate the state $\mathbf x_t$, of the system given the observed field and current estimate of the model parameters to calculate expected log-likelihood. In the M-step we estimate the system parameters by maximising the expected log-likelihood computed in the E-step. The resulting parameters are then used to estimate the distribution of the state vector sequence for the next iteration. The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}.

@@

Since the non-linearity is a sigmoid function first-order Taylor approximation provides a good approximation to the lower bound of the likelihood function, however in general settings with more complicated nonlinear behaviours higher order terms might need to be included, this however introduces intractable computational complexity.
\begin{itemize}
	\item itterative EM and correlation analysis (maybe)
\end{itemize}

>>>>

Here we describe the problem of estimating non-linear state space neural field model using expectation-maximisation (EM) based algorithm \cite{Dempster1977,Shumway2000} is described. EM is an iterative algorithm that converges on the maximum-likelihood estimates of the parameters. Both the E and the M steps of the EM algorithm increase a lower bound on the log-likelihood function of the parameters given the observed field.  In the E-step we estimate the state $\mathbf x_t$, of the system given the observed field and current estimate of the model parameters to calculate expected log-likelihood. In the M-step we estimate the system parameters by maximising the expected log-likelihood computed in the E-step. The resulting parameters are then used to estimate the distribution of the state vector sequence for the next iteration. The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}.

@@

Since the non-linearity is a sigmoid function first-order Taylor approximation provides a good approximation to the lower bound of the likelihood function, however in general settings with more complicated nonlinear behaviours higher order terms might need to be included, this however introduces intractable computational complexity.
\begin{itemize}
	\item itterative EM and correlation analysis (maybe)
\end{itemize}
\scetion{Conclusion}
